---
title: "COOP"
output:
  html_document: default
  pdf_document: default
---


## Articles

https://electrek.co/2020/02/23/tesla-model-3-outsold-other-evs-california-2019-combined/

https://afdc.energy.gov/files/u/publication/ev_emissions_impact.pdf

https://www.bts.gov/content/average-cost-owning-and-operating-automobile 

https://tableau.cnra.ca.gov/t/CNRA_CEC/views/DMVDataPortal_15986380698710/STOCK_Dashboard?%3AshowAppBanner=false&%3Adisplay_count=n&%3AshowVizHome=n&%3Aorigin=viz_share_link&%3AisGuestRedirectFromVizportal=y&%3Aembed=y
 
 
 
## Packages Initialization
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
#install.packages("sqldf")
library(sqldf)
```

## Data Definitions
```{r}
sales = read.csv('data/sales_model_date.csv') #Number of cars sold per model from 2011 - 2019
afv = read.csv('data/AFV_models.csv') # model offerings per Manufacturer from 1991 - 2019
veh = read.csv('data/Vehicles.csv') %>% 
  mutate(City.Fuel.Economy = as.numeric(City.Fuel.Economy),
         Highway.Fuel.Economy = as.numeric(Highway.Fuel.Economy)) # car model specifications
pop = read.csv('data/Electric_Vehicle_Population_Data.csv')

stations = read.csv('alt_fuel_stations_all.csv')
```

```{r}
g1 = pop %>% filter(str_detect(Electric.Vehicle.Type,'(PHEV)')) %>% select(Electric.Range)
g2 = pop %>% filter(str_detect(Electric.Vehicle.Type,'(BEV)')) %>% select(Electric.Range)
```

```{r}
t.test(g1,g2)
```

```{r}
pop %>% group_by(Model) %>% summarise(mean(Electric.Range)) 
```
```{r}
sqldf('select Model, avg("Electric.Range") from pop group by Model')
```


## model with most/least sales in 2019
```{r}
sqldf("select vehicle,max(X2019) from sales") #Tesla Model 3
sqldf("select vehicle,min(X2019) from sales") #Mercedes B-Class (B250e)

```

## How many of each types exist?
```{r}
sqldf("select count(*) from sales where Type = 'EV'") #23
sqldf("select count(*) from sales where Type = 'PHEV'") #32
```

> 23 EVs and 32 PHEVs in the dataset

## Using CTEs, find the max/min counts of vehicles per year

```{r}
sqldf("with t1 as (SELECT 2011 - 1 + ROW_NUMBER() OVER(ORDER BY Vehicle) Year,*
  FROM (select vehicle,max(X2011) as max from sales
      union all select vehicle,max(X2012) from sales
       union all select vehicle,max(X2013) from sales
       union all select vehicle,max(X2014) from sales
       union all select vehicle,max(X2015) from sales
       union all select vehicle,max(X2016) from sales
       union all select vehicle,max(X2017) from sales
       union all select vehicle,max(X2018) from sales
       union all select vehicle,max(X2019) from sales)),
  t2 as (
  SELECT 2011 - 1 + ROW_NUMBER() OVER(ORDER BY Vehicle) Year,*
  FROM (select vehicle,min(X2011) as min from sales
      union all select vehicle,min(X2012) from sales
       union all select vehicle,min(X2013) from sales
       union all select vehicle,min(X2014) from sales
       union all select vehicle,min(X2015) from sales
       union all select vehicle,min(X2016) from sales
       union all select vehicle,min(X2017) from sales
       union all select vehicle,min(X2018) from sales
       union all select vehicle,min(X2019) from sales)
  ) 
  select * from t1 
  inner join t2 on t1.year = t2.year")
```

> Tesla seems to be doing the best in the last 5 years

## Find the 3rd highest count by Model from afv, return null is no 3rd highest

```{r}
sqldf(" select
          ifnull(( select 
                      distinct total 
                   from afv 
                   order by total desc 
                   limit 1 offset 3),null
                   ) as thirdHighest 
          from afv limit 1")
```


```{r}
veh
```

```{r}
sqldf('
      WITH AVG_CITY (Model,avg_cty_econ) AS (
      select
        Model,
        avg("City.Fuel.Economy")
      from veh 
      group by Model
      order by avg("City.Fuel.Economy") DESC),
      AVG_HWY (Model,avg_hwy_econ) AS (
      select 
        Model,
        avg("Highway.Fuel.Economy") 
      from veh 
      group by Model
       order by avg("Highway.Fuel.Economy") DESC)
      select 
        t1.Model, 
        t1.avg_cty_econ,
        t2.avg_hwy_econ 
      from AVG_CITY as t1
      inner join AVG_HWY as t2 on t1.Model = t2.Model
      order by t2.avg_hwy_econ desc
      ')
```

```{r}
veh %>% head(2)
```

## Find all models and unique ids whose "City.Fuel.Economy" is less than its group model's average
```{r}
sqldf('
      with t1 (ID,Year,Model,Avg) as (
      select 
        "Automobile.ID",
        "Model.Year",
        Model, 
        avg("City.Fuel.Economy") as avg_cty
      from veh
      group by Model
      having "City.Fuel.Economy" < avg_cty)
      select * from t1')
```

## Find models who have the lowest "City.Fuel.Economy" per Manufacturer & for Chevy alone

- group by manufacturer, find lowest city econ 

- be aware that two different models from the same manufacturer could have the same lowest value

```{r}
sqldf('select 
        distinct veh.Manufacturer,
        Model,
        "City.Fuel.Economy" as city_econ 
       from veh 
      inner join (
        select 
          Manufacturer,
          min("City.Fuel.Economy") as min 
        from veh 
       
        group by Manufacturer) as ref
      on ref.Manufacturer = veh.Manufacturer and ref.min = veh."City.Fuel.Economy"
      ')
```

### sanity check for to see if above sql returns expected output for Chevy
```{r}
veh %>% select(Manufacturer,Model,City.Fuel.Economy) %>% filter(Manufacturer == "Chevrolet") %>% filter(City.Fuel.Economy == min(City.Fuel.Economy)) %>% distinct()
```

```{r}
veh %>% select(X110v.Charge.Time) %>% unique()
```


## Use case when to bracket models into LOW,MED,HIGH groups based on charging time
```{r}
sqldf('
      select distinct Model, "X110v.Charge.Time",
      case
        when "X110v.Charge.Time" in ("6 hrs","6 hours","6-8 hours") then "LOW"
        when "X110v.Charge.Time" in ("15 hours","20 hours") then "MED"
        when "X110v.Charge.Time" = "22-24 hrs" then "HIGH"
        else NULL
      end as "charge_durs"
      from veh')
```

## Window functions

- ROW_NUMBER() OVER ( ORDER BY COL) AS ...
    - indexes each row by incremented values
    
- LAG(COL,N) OVER (ORDER BY COL) AS ... 
    - provides access to a row at a specified physical offset which comes before current row

```{r}
wf = afv %>% filter(Manufacturer == "Chevrolet") %>% pivot_longer(!Manufacturer) %>% mutate(value = ifelse(is.na(value),0,value)) %>% head(29)
sqldf('select  
        ROW_NUMBER() OVER (ORDER BY Manufacturer) ind,
        name,
        value,
        LAG(value,1) OVER (ORDER BY name) prev_counts
      from wf')
```


```{r}
pop
```



```{r}
stations %>% select(Groups.With.Access.Code) %>% table()
```


```{r}
stations %>% select(Access.Days.Time) %>% table() %>% data.frame() %>% filter(str_detect(.,'24 hours daily')) %>% distinct()
```

```{r}
stations %>% select(Access.Days.Time) %>% unique()
```
```{r}
calvls = read.csv('data/CA_LV123.csv')
calvls
```

```{r}
sqldf('select distinct City,count(*) as count from calvls group by City order by count desc')
```

```{r}
calvls %>% filter(Station.Name == "Nissan of Irvine")
```
```{r}
calvls %>% group_by(City) %>% summarise(c=n()) %>% arrange(-c)
```
```{r}
calvls %>% group_by(Station.Name) %>% summarise(c = n()) %>% arrange(-c)
```

```{r}
calvls %>% group_by(City)
```

